# Installation

Madingley is written in C++, but an R version is also available. Although a little bit less flexible perhaps, it should be more than sufficient four this project. And much more familiar.

```{r setup, include=F}
library(MadingleyR)
```


```{r, eval=F}
library(devtools)
install_github('MadingleyR/MadingleyR', subdir='Package', build_vignettes = T, force = T)
```

I could not get the vignettes to build. It crashed on line 52 in Madinlgey.Rmd with this function call:

```{r, eval=F}
MadingleyR::madingley_version()
```
Error message: ./madingley: /lib/x86_64-linux-gnu/libm.so.6: version `GLIBC_2.29' not found (required by ./madingley)
running command 'cd "/home/NINA.NO/anders.kolstad/R/x86_64-pc-linux-gnu-library/4.0/MadingleyR/lin_exec/" && ./madingley version' had status 1Error in x[[jj]][iseq] <- vjj : replacement has length zero

This tells us that R cannot fint libm versionGLIBC 2.29. The installation works fine on my windows machine, but not on the linux server. 

So, I installed without building vignettes...

```{r}
vignette(package ="MadingleyR") 
```
They can, however, be view [here](https://github.com/MadingleyR/MadingleyR).

But, as long as the model can't find libm, the model cannot run.




