# Step 1 & 2
```{r setup, include=F}
library(MadingleyR)
knitr::opts_chunk$set(comment='%') # the default ## read as headers in the knitted version
```

## Load data

Spatial model domain = c(min_long, max_long, min_lat, max_lat)
```{r,  message=FALSE}
spatial_window = c(31, 35, -5, -1)
```

Plot the spatial window to check selection
```{r}
plot_spatialwindow(spatial_window)
```


Using default data to initialise the model. The spatial inputs is a relatively big list of raster files, but the others are small data frames. 
```{r, message=FALSE, results='hide', cache=T}
sptl_inp = madingley_inputs("spatial inputs")
chrt_def = madingley_inputs("cohort definition")
stck_def = madingley_inputs("stock definition")
mdl_prms = madingley_inputs("model parameters") # useful later for running the model
```


The Spatial Input is the environmental variables in the form of several raster layers. I think the 12 slots are months of a year.
```{r}
str(sptl_inp,2)
```

The cohort definitions look like this, and can be modified by the user. 
```{r}
DT::datatable(chrt_def,
  extensions = "FixedColumns",
  options = list(
    scrollX = TRUE,
    scrollY=T,
    pageLength = 10))
```
The stock definitions are in the default input data limited to evergreen and deciduous plants. 
```{r}
DT::datatable(stck_def,
  extensions = "FixedColumns",
  options = list(
    scrollX = TRUE,
    scrollY=T,
    pageLength = 10))
```
The model parameters are perhaps a bit more cryptic:
```{r}
DT::datatable(mdl_prms,
  extensions = "FixedColumns",
  options = list(
    scrollX = TRUE,
    scrollY=T,
    pageLength = 20))
```
They are divided into 12 categories:
```{r}
cat(unique(mdl_prms$category), sep="\n")
```


## Initialisation

Model initialisation is done using ```madingley_init()``` by supplying the input data above (excl. model parameters). Also one can set the maximum number of cohors allowed per cell. The default is 500 and max is 1000. a higher number increases computation time due to more cohort interactions. 

```{r, cache=T, eval=F}
mdata = madingley_init(spatial_window = spatial_window,
                       cohort_def = chrt_def,
                       stock_def = stck_def,
                       spatial_inputs = sptl_inp,
                       max_cohort = 50)
```
This throws an error:
```./madingley: /lib/x86_64-linux-gnu/libm.so.6: version GLIBC_2.29 not found (required by ./madingley)
cannot open file /tmp/RtmpEF9AGw/madingley_outs_14_12_21_13_39_45/cohort_properties/FullCohortProperties_99999.csv: No such file or directoryError in file(file, "rt") : cannot open the connection```



