# Step 1 - 3
```{r setup, include=F}
library(MadingleyR)
knitr::opts_chunk$set(comment='%') # the default ## read as headers in the knitted version
```



## Step 1 - Load data

Defining the study area in degree long-lat.
```{r,  message=FALSE}
spatial_window = c(31, 35, -5, -1)
```

Plot the spatial window to check selection
```{r}
plot_spatialwindow(spatial_window)
```
This corresponds to somewhere in the Serengetti I think.

This is the  default data to initialise the model. The spatial inputs is a relatively big list of raster files, but the others are small data frames. 
```{r, message=FALSE, results='hide', cache=T}
sptl_inp = madingley_inputs("spatial inputs")
chrt_def = madingley_inputs("cohort definition")
stck_def = madingley_inputs("stock definition")
mdl_prms = madingley_inputs("model parameters") # useful later for running the model
```


The Spatial Input is the environmental variables in the form of several raster layers. I think the 12 slots are months of a year.
```{r}
str(sptl_inp,2)
```

```{r}
temp <- sptl_inp$precipitation
class(temp)
```
```{r}
temp
```
The resolution is 1 degree, which is approx. 111 km at the equator. 

Plotting monsth jan-april
```{r}
plot(temp[[1:4]])
rm(temp)
```
The data has global coverage.


The cohort definitions look like this, and can be modified by the user. 
```{r}
DT::datatable(chrt_def,
  extensions = "FixedColumns",
  options = list(
    scrollX = TRUE,
    scrollY=T,
    pageLength = 10))
```
The stock definitions are in the default input data limited to evergreen and deciduous plants. 
```{r}
DT::datatable(stck_def,
  extensions = "FixedColumns",
  options = list(
    scrollX = TRUE,
    scrollY=T,
    pageLength = 10))
```
The model parameters are perhaps a bit more cryptic:
```{r}
DT::datatable(mdl_prms,
  extensions = "FixedColumns",
  options = list(
    scrollX = TRUE,
    scrollY=T,
    pageLength = 20))
```
They are divided into 12 categories:
```{r}
cat(unique(mdl_prms$category), sep="\n")
```
## Step 2 - Modify input parameters
We will not do any modifications at this stage

## Step 3 - Initialisation

Model initialisation is done using ```madingley_init()``` by supplying the input data above (excl. model parameters). Also one can set the maximum number of cohors allowed per cell. The default is 500 and max is 1000. a higher number increases computation time due to more cohort interactions. 

```{r, cache=T, eval=T}
mdata <-  madingley_init(spatial_window = spatial_window,
                       cohort_def = chrt_def,
                       stock_def = stck_def,
                       spatial_inputs = sptl_inp,
                       max_cohort = 100)
```
Runtime is just a few seconds.

```{r}
str(mdata, 1)
```

The output contain the cohors and stocks for each pixel.

```{r}
save(mdata, chrt_def, mdl_prms, sptl_inp, stck_def, 
        file = "data/initOut_demo.RData")
```


