# Step 4 - Spin-up and demonstrating plotting functions

```{r}
library(MadingleyR)
knitr::opts_chunk$set(comment='%')
load("data/initOut_demo.RData")
```

The spin-up usualy requires a 100-1000 year simulation for the ecosystem parameters to stabilise. Here we will use just 2 years for demonstration. The function takes as input the output data fram ```madingley_init()``` as well as the standard model input variables (which can be modified). 

The below code is the same as in the vignette example except I had to define my own output folder for some reason. 
```{r, cache=T}
mdata2 <-  madingley_run(
  out_dir = "temp",
  madingley_data = mdata, 
  years = 2, 
  cohort_def = chrt_def, 
  stock_def = stck_def, 
  spatial_inputs = sptl_inp, 
  model_parameters = mdl_prms,
  max_cohort = 100)

# Runtime approx 30 sec

# View the contents of mdata2
str(mdata2,1)
```

This output is similar to the previous outpur from madingley_init() only it also contains stocks and cohorts as timeslines. 

Lets view these timelines.
```{r, fig.height=7, fig.width=7, message=FALSE,error=FALSE,linewidth=90}
plot_timelines(mdata2)
```

Plot histograms for body mass:
```{r, fig.height=7, fig.width=7, message=FALSE,error=FALSE,linewidth=90}
plot_densities(mdata2)
```

Plot trophic pyramid
```{r, fig.height=7, fig.width=7, message=FALSE,error=FALSE,linewidth=90}
plot_trophicpyramid(mdata2)
```

Create log10-binned food-web plot
```{r, fig.height=6, fig.width=7, message=FALSE,error=FALSE,linewidth=90}
plot_foodweb(mdata2, max_flows = 5)
```

Plot the spatial biomass
```{r, fig.height=7, fig.width=9, message=FALSE,error=FALSE,linewidth=90}
plot_spatialbiomass(mdata2, functional_filter = TRUE)
```

```{r}
save(mdata2, chrt_def, mdl_prms, sptl_inp, stck_def, 
        file = "data/spinupOut_demo.RData")
```


