# Trophic pyramids

Exploring the biomass time series produced from running a 10 year spinup for 16 grid cells in central Norway.

This page was last updated ```r Sys.time() ```

```{r setup, include=F}
library(MadingleyR)
library(ggplot2)
library(ggpubr)

knitr::opts_chunk$set(comment='%')
load("data/NorwayDefaultSpinup.RData")
```

```{r, fig.height=5, fig.width=7, message=FALSE,error=FALSE,linewidth=90}
plot_trophicpyramid(mdata2)
```

The autotrophic or plant biomass is 10^15.17 kg. Written in a more familiar different way, that is the same as
```{r}
((10^15.17)/1000)/10^9 # divided by 1 000 to go from kg to tons, and 1 000 000 000 to get Gt
```
Gigatons (Gt).

Where does this number come from? 
I can recreate the number by summing the biomass for the whole last year of the simulation:
```{r}
biom <- sum(mdata2$time_line_stocks$TotalStockBiomass[
  mdata2$time_line_stocks$Year==max(mdata2$time_line_stocks$Year) ])

#Converted to base10 scientific
paste0("10^", round(
  log10(biom), 2))

```

According to [Wikipedia](https://en.wikipedia.org/wiki/Biomass_(ecology)) the total biomass of the world is about 1 000 billion tons (assuming 50% C). This is the same as 1 000 gigatons Gt. A [different source](https://www.weforum.org/agenda/2021/08/total-biomass-weight-species-earth) says 500 Gt for the total world biomass.

Our estimate for the plant biomass of south Norway is therefore about 50% bigger than the total biomass of the planet! 

[Hoeks et al 2020 Glo. Eco. Bio.](https://onlinelibrary.wiley.com/doi/full/10.1111/geb.13354) got comparable values to us when using the same R package and same study extent (4x4 degrees). If our numbers are wrong in thi example, then their numbers must be wrong as well.

[Harfoot et al 2014](https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1001841) got much lower values using the C# version of Madingley. For a single grid cell (aseasonal ecosystem in tropical Uganda) they got 24 611 tons of plant biomass. If we multiply that with 16 to become comparable to our study extent, and convert it to Gt, we get 

```{r, echo=F}
(24611*16) / 10^9
```
This is much, much lower than our estimate.
 
This is roughly the script that produces the trophic pyramid figure above (plot_trophicpyramid())
```{r, eval=F}
years <-  0
tl <-  mdata2$time_line_stocks
tl <-  aggregate(tl, by = list(tl$Year), FUN = sum)
tl <-  tl[(nrow(tl) - length(years) + 1):nrow(tl), ]
(biom <- mean(tl$TotalStockBiomass))
```
I suspect that the unit in mdata2$time_line_stocks is grams, and not kg. As we will see below, it's grams for the heterotrofic cohorts.
But even so, the numbers are way off.
 
How do our monthly estimates compare to Harfoot et al 2014.
This is our maximum monthly estimate in the last year of the spin up, converted to Gt:
```{r}
(temp <- max(
  mdata2$time_line_stocks$TotalStockBiomass[
   mdata2$time_line_stocks$Year==max(mdata2$time_line_stocks$Year) ])/1000/10^9)

```
This estimate is still too big. 
Is we assume the the unit should've been g and not kg, we still get 0.2 Gt (compare with 0.0004 in Harfoot et l 2014).

There must be more to this than the g and kg mistage and the fact that the biomass is an accumulated monthly sum.  

But why does the R function use an accumulated monthly sum? I can not think of a good reason.

It makes more sense to me to just get the standing biomass for a given month, perhaps the maximum monthly value. In our area in Norway the plant biomass peaks in September.
```{r, echo=F}
biom <- mdata2$time_line_stocks$TotalStockBiomass[
  mdata2$time_line_stocks$Year==max(mdata2$time_line_stocks$Year) ]

myY <- seq(0,max(biom), length.out = 5)
myY2 <- round(seq(0,log10(max(biom)), length.out = 5), 0)
myY2 <- paste0("10^", myY2)


plot(biom, type="b", axes=F, ylab="Plant biomass (kg)", xlab="Month")
axis(1, 1:12, month.abb, las=2)
axis(2, myY, myY2)
```



What about the herbivore data. How are those biomass estimates that are reported in the trophic pyramid above calculated?
In the trophic pyramid we see a value of about 0.09 Gt:
```{r}
((10^10.97)/1000)/10^9 # divided by 1000 to go from kg to tons, and 1 000 000 000 to get Gt
```
Harfoot et al 2014 report for their study system (converted to 16 grid cells and to Gt)

```{r}
(240*16)/10^9
```

We have a similar situation here. 

Here is the heterotrophic biomass over the last year of our simulation

```{r}
dat <- mdata2$time_line_cohorts

dat <- dat[dat$Year == max(dat$Year),] #subset to only the last year

datm <- data.table::melt(dat,
                         id.vars = c("Year", "Month"),
                         variable.name = "FG",
                         value.name = "biomass")

datm$tons <- datm$biomass/1000000  # assuming the unit is g, so i divide by 10^6 to give tons
ggplot(dat = datm, aes(x = Month, y = tons, 
                       colour = FG,
                       linetype=FG))+
  geom_line(size=1.5)+
  theme_bw(base_size = 12)+
  ylab("Biomass (tons)")
```

Now, let me try to recreate the number in the trophic pyramid.
Combining the herbivores
```{r}
datherb <- datm[datm$FG == "Biomass_FG_0" |
                  datm$FG == "Biomass_FG_3" |
                  datm$FG == "Biomass_FG_6" ,]
```

Summed biomass across all months in the raw/unknown units
```{r}
sum(datherb$biomass)
```
If we assume that units are in grams, this is the same as
```{r}
sum(datherb$biomass)/1000/10^9
```
Gt, which is quite close the number we were aiming for. 

It may look like the heterotrophic biomass and the autrophic biomass is calculated in the same way.

The ration between autrophic and herbivore biomass is in our Norwegian case study
```{r}
(10^10.97)/(10^15.17)
```
, or 0.000063,

and in Harfoot et al 2014 it is 
```{r}
240.2/24610.8
```
Norway apparently has a much higher autrophic biomass compared to heterotrophic biomass, but I don't think that is the reality

